Our Title

### 2015-05-01 draft.tex
Fixed random references, and bad apostrophes.

Since ISIL began to seize power and land in Iraq and Syria, many of studied what exactly drives the organization. Perhaps the most complete discussion of the group's motivation comes from Graeme Woods’ article \enquote{What ISIL Really Wants,} which appeared in the March 2015 issue of The Atlantic \cite{Wood2015}. In assessing ISIL and its goals, it is first necessary to recognize its differences with other militant Islamic terrorist organizations.

@@

Much discussion has surrounded the flow of foreign fighters into ISIL controlled territory. Again, this is rooted in the groups’ religious beliefs. According to Salafism, once a caliphate is established, all true believers must move to it, unless completely unable to do so.

@@

In assessing how ISIL may acquire a nuclear weapon, it is necessary to consider potentially helpful alliances and allies, as well as their enemies.  As with most things related to ISIL and the Middle East, this is not necessarily as straightforward as one would assume.  While some relationships are fixed points of reference, most are dynamic and characterized by opportunism, both on the part of ISIL and the organization or state who seeks to gain some benefit from supporting or opposing them.  The following discussion is not intended to be all encompassing but instead will focus on key players and how they may help or hinder ISIL’s nuclear ambitions.

At first glance, the strategic landscape does not look favorable to ISIL as depicted in \autoref{fig:ISILRelationships} below.  While somewhat dated, it serves a decent representation of some of the key groups and states that are in direct conflict and collaboration with ISIL.  ISIL’s key relationships can be broadly grouped into three categories:  countries and groups at war with ISIL, former or potential future allies, and groups that support ISIL.

@@

Additionally, there are longstanding mistrust, perceptions, and illicit alliances that could cause large fissures  without the help of ISIL.  While the Iraqi government is cooperating with the U.S. in battling ISIL, some clerics and government officials have publicly claimed the CIA was responsible for the rise of ISIL \cite{Kirkpatrick2014}.  Whether this is true is largely irrelevant if the mere perception affects the extent of U.S. - Iraqi cooperation against ISIL, especially if the need would arise to place U.S. troops in combat.  Other reports have the Turkey military providing safe travel and weapons for ISIL fighters attacking Kurdish forces, which have been a thorn in the side of Turkey \cite{Guiton2014}.  Finally, it appears that wealthy Saudis, who have a history of funding terrorist organizations, are continuing to provide ISIL with funding \cite{Windrem2014}. While this funding and free travel is not a large contributor to ISIL’s overall success, it represents larger issues faced in isolating ISIL from the international community and shows a willingness of certain groups to deal with ISIL if they see personal benefit despite the international condemnation.  Currently, it is just guns, cash, and ammunition, but as time wears on does someone sell nuclear components or technology to ISIL in an attempt to accomplish some goal?  It may sound unfathomable, but perhaps no more irrational that the U.S. arming militias and rebels in Syria, despite the fact that the current issues facing the Middle East with regards to terrorism and ISIL can be traced to the U.S. arming and training of Afghani rebels in the 1980s.

@@

Although the increasing coalition arrayed against ISIL is a promising curb to their  nuclear ambitions, the growth of ISIL’s alliances is not.  Currently, ISIL has 32 affiliates throughout the world of varying size and level of support \cite{IntelCenter2015}.  A map of the countries where those organizations operate is shown in \autoref{fig:affmap}.  Two important countries in the region with respect to nuclear issues stand out.  First, Iran is currently opposed to ISIL, has no organizations pledging support to ISIL, and is funneling support for Iraq and Hezbollah to combat the spread of ISIL.  However, Iran is no stranger to supporting terrorist organizations that it believes can advance its interests, and it has previously shown tendencies to cross sectarian lines for even tenuous, direct, or indirect gain to Iran \cite{pollack2014unthinkable}.  Pakistan, on the other hand, has had no shortage of declared support, and many believe that trend is to continue \cite{IntelCenter2015,Walsh2014}.  While Pakistan is firm in statements about the control of its nuclear weapons, their program has had a history of proliferation and is widely considered to have ongoing commitments to sell nuclear weapons should the need arise \cite{Langewiesche2005,Henderson2015,Smith2011}.  Due to the consequences, the low probability event of Pakistan or Iran aiding ISIL’s nuclear ambitions poses unique challenges in the goal of preventing ISIL from obtaining nuclear weapons.

@@

When considering ISIL’s nuclear ambitions, one country not currently entangled in the ISIL conflict must be considered: North Korea.  In the past, North Korea has shown a willingness to export nuclear technology \cite{Wit2013}.  Some have argued the test of an uranium device in 2013 was an advertisement to the world that it could produce excess nuclear weapons that may be available for sale \cite{Allison2013}.  Currently, there is no known relationship between ISIL and North Korea.  However, given the unpredictability of the North Korean regime, coupled with the fact that there are operatives throughout ISIL controlled territory that would have North Korean contacts due to the prior Syrian - North Korean nuclear collaborations, any attempt to stop ISIL from obtaining a nuclear weapon needs to account for this potential pathway.

@@

The only true difference in the process of producing uranium for use in a nuclear reactor or in a weapon is the amount of enrichment carried out. Most enrichment is done using gas centrifuges, however this require significant infrastructure that is almost definitely beyond the capabilities of ISIL \cite{Benedict1981}.  Furthermore even if ISIL were able to gain control over an enrichment facility, the length of time it would take them to enrich the uranium for a weapon would likely be long enough to allow other countries to intervene militarily. For example, if ISIL were to gain control of Iran’s Natanz facility, it would take them almost two months to enrich the uranium necessary for a weapon. During that time, there would be obvious tell-tail signs of ISIL attempting to enrich uranium, such as power at that site, without needing to monitor nuclear signatures. Thus the most significant piece of stopping ISIL from enriching uranium for a weapon is simply identifying enrichment facilities, which are difficult to hide, with in its territory.

While ISIL could theoretically use laser separation to enrich uranium, which is significantly more difficult to detect, this is very unlikely. Currently, there is only one prototype process under investigation and it’s in Australia \cite{Moody2014}.  Thus, it seems unlikely that ISIL would be able to use this method to enrich uranium, and thus it does not need to be monitored for.

@@

While uranium appears in nature, plutonium is only available through irradiation in a nuclear reactor \cite{Benedict1981}. It is formed when \ce{^{238}U} undergoes the process known as neutron capture \cite{Benedict1981}.  After a double beta decay, \ce{^{239}Pu}, which can be used in nuclear weapons, is produced. [4] While heavy water reactors do have a higher percentage of \ce{^{238}U}, and thus form \ce{^{239}Pu} at a faster rate, it is possible to also use a light water reactor to produce plutonium \cite{Moody2014}.  While ISIL could theoretically gain control of a commercial nuclear reactor for plutonium production, this seems incredibly unlikely. The only commercial nuclear reactors currently near their territory are located in Iran. [5] Furthermore, while countries like Saudi Arabia, Egypt, Turkey, and the UAE are all moving towards nuclear power, it will be many years before any of those countries have commercial plants. [5] While there are not many commercial reactors that ISIL could easily use to produce plutonium, there are many research reactors scattered throughout the Middle East. [6] However, ISIL would need to make very significant territory gains for any of the closest research reactors to be in danger. [7] Thus, it seems unlikely that ISIL would be able to gain control of a nuclear reactor for plutonium production.

It should be noted that Syria may have had a reactor for the production of plutonium in the past. However, that site was bombed and destroyed by an Israeli airstrike in 2007. [6] It is not impossible that ISIL could obtain the plans for that design. That being said, construction on a nuclear power plant typically takes at least 4 years. [8] Furthermore, nuclear reactors are incredibly difficult to hide, as seen by Israel’s 2007 air raid against Syria. [6] Thus, it seems unlikely that ISIL would be able to complete such a reactor. Therefore, there is a very low probability that ISIL could covertly produce it’s own plutonium, and thus little need to monitor for reprocessing signals from ISIL territory.

\subsection{??????  Add in these refs    ???????}

% [3] “Iran’s Nuclear Timetable,” Wisconisn Projct on Nuclear Arms Control, < http://www.iranwatch.org/our-publications/articles-reports/irans-nuclear-timetable#i>, 24 February, 2015.
% [4] “Nuclear Fission Energy,” Lawrence Berkeley National Laboratory, http://www2.lbl.gov/abc/wallchart/chapters/14/1.html, Accessed: 23 April, 2015.
% [5] “Nuclear Power in the World Today,” World Nuclear Association, <http://www.world-nuclear.org/info/Current-and-Future-Generation/Nuclear-Power-in-the-World-Today/>, February, 2015.
% [6] “Emerging Nuclear Energy Countries,” World Nuclear Association, <http://www.world-nuclear.org/info/Country-Profiles/Others/Emerging-Nuclear-Energy-Countries/>, March 2015.
% [7] “Battle for Iraq and Syria in maps,” BBC, <http://www.bbc.com/news/world-middle-east-27838034>, 24 April, 2015.
% [8] “Building New Nuclear Facilities,” Nuclear Energy Institute, <http://www.nei.org/Issues-Policy/New-Nuclear-Energy-Facilities/Building-New-Nuclear-Facilities>, 2015.

@@

\subsection{???????? Scott's added brief RDD info, needs content proofing. Andrew tried to use it to link the above section to the case study    ?????????}

@@

From the U.S. perspective, 9/11 serves as a case study to understand the shortcomings of IC, subsequent recommended changes, and possible shortcomings facing the IC with respect to ISIL. The 9/11 Report found that there are cultural and organizational factors that hinder strategic communication between U.S intelligence agencies (See \autoref{app:partners} for a disucussion on strategic communication on the international level, which also applies at the interagency level).  Prior to 9/11, U.S intelligence agencies operated in a culture dominated by \enquote{need-to-know}, which implied that organizations were given discretion in sharing data \cite{McConnell2008}. In the context of a dynamic ISIL threat, this means that agencies often have static analyses formed from an incomplete picture and lacking a competing analysis derived from the same data. The divisions between individual agencies render them incapable of accessing, prioritizing, translating, analyzing, and relaying the information for the policymakers to respond to in a timely manner. As a result, information flow is hindered by each agency's networks and databases lacking protocols to interact with other agencies' networks and databases.  Organizational inertia, standard operating procedures, and guidelines hinder the ability to change or implement new ideas. For example, the CIA’s culture of risk aversion in the 1990s was detrimental to its intelligence capabilities and prevented them from properly assessing and taking action against the rising threat of terrorism \cite{Zegart2005}.

The situation is only complicated when considering international intelligence agencies and the sharing of data. To simplify the problem, the factors considered in determining the level of cooperation and the nature of an intelligence partnership between U.S. and other countries were distilled down to the historical relationships between both parties, the country’s relationship to ISIL, its military and intelligence capability, and its vulnerability and geographical proximity to ISIL. The role of international entities is gauged by the organization’s functions, credibility, and past success in relaying information and promptly responding to crises.  These factors are further described in \autoref{app:intel}.

@@

\subsection{??????   Add this reference in   ?????}
[3] JUSTINE DRENNAN, “Who Has Contributed What in the Coalition Against the Islamic State?”, NOVEMBER 12, 2014

More than 60 countries have joined the \enquote{global coalition to degrade and defeat ISIL.} [3] The member countries can choose their method and level of participation. Saudi Arabia trains and equips Syrian opposition forces. Jordan’s key role will be intelligence acquiring and sharing. Germany, France, and the United Kingdom are conducting airstrikes.

The number of countries in the coalition is growing with Obama’s campaign, but the bar for inclusion is fairly low. While many countries have committed military capabilities to combatting ISIL, some countries join the coalition by merely declaring their intent to contribute. All these member countries don’t share the same understanding of whom the real enemy is - ISIL or the Syrian regime led by President Bashar al-Assad. For example, Turkey has urged the coalition to topple the Assad regime as a condition to provide its air bases to carry out attacks against ISIL.

@@

UK and France have been U.S.’ allies even since the country launched its global fight against terrorism after the 9/11 incident. Hence, the U.S. can expect UK specifically to be its closest ally in preventing ISIL’ nuclear terrorism effort. The two countries \enquote{special relationship} dated back to Churchillian era means that the U.S. can usually count on the UK for mutual support, especially considering ISIL’ threats to the country is very high. UK has been ISIL’ main target for terrorist attacks, and given UK’s nuclear capability, the country is extra vulnerable to terrorist attacks on any of its nuclear installations. So far, UK has joined U.S.-led coalition against ISIL, nevertheless, prospect for bilateral cooperation between the two countries cannot be disregarded.

Meanwhile, France is U.S.’ oldest ally and relationship remains active and friendly. France supported U.S. in various issues, however, France was skeptical when U.S. invaded Iraq after the 9/11 incident. ISIL has also posed great threats to France (being a successful nuclear energy producer and all), France can be vulnerable to attacks on its nuclear energy facilities as well as its possession of nuclear weapons stocks. France is also among ISIL’ main recruitments considering France high marginalized and discontented Muslim populations. France’s high ISIL threats and vulnerabilities made the country a good collaborator in effectively fighting ISIL.

Nuclear capabilities in the hand of weak or unstable government give opportunity for ISIL to acquire nuclear capabilities. Nuclear capabilities of Pakistan and its weak governance as well as security vulnerabilities is of concern to the U.S.. Looking at U.S.-Pakistan relationship, it has its ups and down. Over the decades, Pakistan has been a great ally to the U.S., but there are a few instance where the two countries were in disagreement. U.S. viewed Pakistan as its ally and threat in the country’s global war on terror, in which Pakistan is seen as the ‘weakest link’ as it continue to fail in containing act of terrorism going on in its own country. U.S. need to prop Pakistan and reduce the country’s vulnerabilities of ISIL penetrations and the possibility of ISIL acquiring Pakistan’s nuclear weapons.

India is also another country with nuclear weapon capability, and is mostly on the \enquote{other side} of U.S.’ security policies. India, together with Russia was among the pioneers of Non-Aligned Movement (NAM), in which views on security issues tend to be divergent with the U.S.. However, as of recently, the U.S. and India have set up ‘global strategic partnership’, based on shared democratic values and increasing convergence of interests on bilateral, regional and global issues; key recent developments in the relationship include the rapid growth of India's economy and bilateral trade, the close links between the Indian and American computer and Internet industries, a geopolitical coalition to balance the rise of China, the weakening of U.S.-Pakistan relations, among others. U.S. has cooperated with India in area of civil nuclear, and this can be expanded to development of nuclear security hopefully in the future. However, unlike Pakistan, India believed ISIL’ threats as periphery issue to the country.

In analyzing Russia as a possible bilateral partner, the two country has had strained relationship, increasingly since the economic sanctions imposed by the U.S. to Russia. Dispute between the two countries have centered on power struggle for new world order, and recently exacerbated by Russia’s annexation of Crimea. The two countries have been historical rival and the Cold War has proven to be one of the expansive and stretched out power struggle between the two. However, both countries realized and agreed to maintain dialogue in several areas including terrorism. Being misaligned with Russia on ISIL and nuclear terrorism issue could be serious mistakes as Russia’s nuclear capability, if not protected and are allowed to be shared in the wrong hand, could have grave catastrophic impact to global security environment.

China, another U.S.’ partner and rival at times, also have nuclear capabilities. China and the U.S. has close partnership particularly in economic sector, two-way trade between the two countries is over U.S.\$590 billion in 2014 alone. Chinese Muslim Uyghur militants have been associated with ISIL and efforts to undertake terrorist attacks in China. This strengthen the need for collaboration between China and U.S.; to ensure any ISIL activities can be contained in China and the country’s nuclear weapons is protected from any probability of ISIL acquiring them.

@@

Iran and U.S. shared a common enemy factor in terms of ISIL. As the nuclear talks between the two countries are being hashed out, both countries have committed on fighting ISIL together. This can be a good opportunity for U.S. to combat ISIL more effectively as U.S. cannot disregard Iran’s influence and role in Iraq’s security landscape and their fight against ISIL.
Meanwhile, collaborating with Israel is considered a gray area and not clear-cut as it seems. Israel has been known to voice their support of ISIL, keeping in line with ISIL’s common enemy with Israel i.e. Iran. As U.S. and Israel have been a strong ally in the Middle East region, this fact presented a conundrum to the U.S..

The U.S. may find other allies in the Middle East region, however, the region is deeply divided in their positions against the ISIL. Although most Middle Eastern countries sees ISIL as a serious threat to the region and its people, some still supported ISIL, if not openly but explicitly. U.S. has worked closely with Jordan, with the U.S. creating a U.S.\$5 billion counterterrorism partnership fund with Jordan and other countries (Libya, Lebanon, Syria) that are considered on the \enquote{frontlines} of U.S.’ fights against ISIL.

The U.S. can find allies in European countries as well, as the region is facing \enquote{recruitment} threats from ISIL. Thousands of Europeans have been recruited by ISIL to fight jihad cause in ISIL’ territories. ISIL is also setting up bases in a number of European countries, expanding its influence beyond the Middle East region. The U.S. can identify these countries and create cooperation efforts in order to effectively eliminate ISIL.

@@

Passive signals for nuclear materials include the gamma ray’s that are given off during natural radioactive decay of the material, and neutrons that result either from spontaneous fission or (\(\alpha\),n) nuclear reactions (the \(\alpha\) particle comes from natural \(\alpha\) decay). The two most likely and common types of material that could be used for a nuclear weapon are highly enrich \ce{^{235}U} and highly enriched \ce{^{239}Pu}, and so these materials will be considered for passive detection. Weapons grade Uranium has a very low rate of released neutrons, and so detection by gamma rays would be recommended. Weapons grade Plutonium releases both, a large amount of neutrons and gamma rays and so both methods of detection are possible.

@@

Of the radiation types mentioned: alpha, beta, and gamma, the gamma emitters are by far the farthest penetrating, most energetic, radioisotope emitters listed.  The shielding suggested will merely attenuate the energy levels of the gamma emissions, but in essence the emission passes through the shielding and continues on its travel through the medium. \autoref{fig:shielding_reqs} shows relative stopping distances in materials for given radiation types. Alpha’s although they can be very energetic upon initial release from the source, are very large particles in comparison to the surroundings.  It is more likely that the 2+ charged alpha ion will interact with matter than the other radiation types, even as depicted a piece of paper is sufficient shielding material to attenuate alpha emissions.

@@

There are three approaches to disrupting ISIL’s operations that will be particularly important in ensuring that ISIL does not acquire a nuclear device:

@@

ISIL’ principal source of income is derived from control and sale of oil, which brings in about \$1 million a day [1]. Additional funds come from kidnap for ransom, extortion, criminal activities, and a small amount from external donations.

@@

A new report by Financial Action Task Force (an international body) finds that it is unclear whether ISIL’ revenue collection will be sustainable over time, and that ISIL will have to gain more territory to remain financially viable \cite{TheEditorialBoard2015,Report2015}.

@@

In particular, sanctions will be levied against ISIL leadership, supporters, and financial facilitators. Effective use of sanctions against the CFO-like figures who manage ISIL funding networks will frustrate ISIL’ ability to raise money and recruit fighters \cite{Cohen2014}.

>>>>

Since ISIL began to seize power and land in Iraq and Syria, many of studied what exactly drives the organization. Perhaps the most complete discussion of the group's motivation comes from Graeme Woods' article \enquote{What ISIL Really Wants,} which appeared in the March 2015 issue of The Atlantic \cite{Wood2015}. In assessing ISIL and its goals, it is first necessary to recognize its differences with other militant Islamic terrorist organizations.

@@

Much discussion has surrounded the flow of foreign fighters into ISIL controlled territory. Again, this is rooted in the groups' religious beliefs. According to Salafism, once a caliphate is established, all true believers must move to it, unless completely unable to do so.

@@

In assessing how ISIL may acquire a nuclear weapon, it is necessary to consider potentially helpful alliances and allies, as well as their enemies.  As with most things related to ISIL and the Middle East, this is not necessarily as straightforward as one would assume.  While some relationships are fixed points of reference, most are dynamic and characterized by opportunism, both on the part of ISIL and the organization or state who seeks to gain some benefit from supporting or opposing them.  The following discussion is not intended to be all encompassing but instead will focus on key players and how they may help or hinder ISIL's nuclear ambitions.

At first glance, the strategic landscape does not look favorable to ISIL as depicted in \autoref{fig:ISILRelationships} below.  While somewhat dated, it serves a decent representation of some of the key groups and states that are in direct conflict and collaboration with ISIL.  ISIL's key relationships can be broadly grouped into three categories:  countries and groups at war with ISIL, former or potential future allies, and groups that support ISIL.

@@

Additionally, there are longstanding mistrust, perceptions, and illicit alliances that could cause large fissures  without the help of ISIL.  While the Iraqi government is cooperating with the U.S. in battling ISIL, some clerics and government officials have publicly claimed the CIA was responsible for the rise of ISIL \cite{Kirkpatrick2014}.  Whether this is true is largely irrelevant if the mere perception affects the extent of U.S. - Iraqi cooperation against ISIL, especially if the need would arise to place U.S. troops in combat.  Other reports have the Turkey military providing safe travel and weapons for ISIL fighters attacking Kurdish forces, which have been a thorn in the side of Turkey \cite{Guiton2014}.  Finally, it appears that wealthy Saudis, who have a history of funding terrorist organizations, are continuing to provide ISIL with funding \cite{Windrem2014}. While this funding and free travel is not a large contributor to ISIL's overall success, it represents larger issues faced in isolating ISIL from the international community and shows a willingness of certain groups to deal with ISIL if they see personal benefit despite the international condemnation.  Currently, it is just guns, cash, and ammunition, but as time wears on does someone sell nuclear components or technology to ISIL in an attempt to accomplish some goal?  It may sound unfathomable, but perhaps no more irrational that the U.S. arming militias and rebels in Syria, despite the fact that the current issues facing the Middle East with regards to terrorism and ISIL can be traced to the U.S. arming and training of Afghani rebels in the 1980s.

@@

Although the increasing coalition arrayed against ISIL is a promising curb to their  nuclear ambitions, the growth of ISIL's alliances is not.  Currently, ISIL has 32 affiliates throughout the world of varying size and level of support \cite{IntelCenter2015}.  A map of the countries where those organizations operate is shown in \autoref{fig:affmap}.  Two important countries in the region with respect to nuclear issues stand out.  First, Iran is currently opposed to ISIL, has no organizations pledging support to ISIL, and is funneling support for Iraq and Hezbollah to combat the spread of ISIL.  However, Iran is no stranger to supporting terrorist organizations that it believes can advance its interests, and it has previously shown tendencies to cross sectarian lines for even tenuous, direct, or indirect gain to Iran \cite{pollack2014unthinkable}.  Pakistan, on the other hand, has had no shortage of declared support, and many believe that trend is to continue \cite{IntelCenter2015,Walsh2014}.  While Pakistan is firm in statements about the control of its nuclear weapons, their program has had a history of proliferation and is widely considered to have ongoing commitments to sell nuclear weapons should the need arise \cite{Langewiesche2005,Henderson2015,Smith2011}.  Due to the consequences, the low probability event of Pakistan or Iran aiding ISIL's nuclear ambitions poses unique challenges in the goal of preventing ISIL from obtaining nuclear weapons.

@@

When considering ISIL's nuclear ambitions, one country not currently entangled in the ISIL conflict must be considered: North Korea.  In the past, North Korea has shown a willingness to export nuclear technology \cite{Wit2013}.  Some have argued the test of an uranium device in 2013 was an advertisement to the world that it could produce excess nuclear weapons that may be available for sale \cite{Allison2013}.  Currently, there is no known relationship between ISIL and North Korea.  However, given the unpredictability of the North Korean regime, coupled with the fact that there are operatives throughout ISIL controlled territory that would have North Korean contacts due to the prior Syrian - North Korean nuclear collaborations, any attempt to stop ISIL from obtaining a nuclear weapon needs to account for this potential pathway.

@@

The only true difference in the process of producing uranium for use in a nuclear reactor or in a weapon is the amount of enrichment carried out. Most enrichment is done using gas centrifuges, however this require significant infrastructure that is almost definitely beyond the capabilities of ISIL \cite{Benedict1981}.  Furthermore even if ISIL were able to gain control over an enrichment facility, the length of time it would take them to enrich the uranium for a weapon would likely be long enough to allow other countries to intervene militarily. For example, if ISIL were to gain control of Iran's Natanz facility, it would take them almost two months to enrich the uranium necessary for a weapon \cite{WisconsinProjectonNuclearArmsControl2015}. During that time, there would be obvious tell-tail signs of ISIL attempting to enrich uranium, such as power at that site, without needing to monitor nuclear signatures. Thus the most significant piece of stopping ISIL from enriching uranium for a weapon is simply identifying enrichment facilities, which are difficult to hide, with in its territory.

While ISIL could theoretically use laser separation to enrich uranium, which is significantly more difficult to detect, this is very unlikely. Currently, there is only one prototype process under investigation and it's in Australia \cite{Moody2014}.  Thus, it seems unlikely that ISIL would be able to use this method to enrich uranium, and thus it does not need to be monitored for.

@@

While uranium appears in nature, plutonium is only available through irradiation in a nuclear reactor \cite{Benedict1981}. It is formed when \ce{^{238}U} undergoes the process known as neutron capture \cite{Benedict1981}.  After a double beta decay, \ce{^{239}Pu}, which can be used in nuclear weapons, is produced \cite{Duderstadt1976}.  While heavy water reactors do have a higher percentage of \ce{^{238}U}, and thus form \ce{^{239}Pu} at a faster rate, it is possible to also use a light water reactor to produce plutonium \cite{Moody2014}.  While ISIL could theoretically gain control of a commercial nuclear reactor for plutonium production, this seems incredibly unlikely. The only commercial nuclear reactors currently near their territory are located in Iran \cite{WorldNuclearAssociation2015}.  Furthermore, while countries like Saudi Arabia, Egypt, Turkey, and the UAE are all moving towards nuclear power, it will be many years before any of those countries have commercial plants \cite{WorldNuclearAssociation2015}.  While there are not many commercial reactors that ISIL could easily use to produce plutonium, there are many research reactors scattered throughout the Middle East \cite{WorldNuclearAssociation2015a}.  However, ISIL would need to make very significant territory gains for any of the closest research reactors to be in danger \cite{BBC2015}.  Thus, it seems unlikely that ISIL would be able to gain control of a nuclear reactor for plutonium production.

It should be noted that Syria may have had a reactor for the production of plutonium in the past. However, that site was bombed and destroyed by an Israeli airstrike in 2007 \cite{WorldNuclearAssociation2015a}. It is not impossible that ISIL could obtain the plans for that design. That being said, construction on a nuclear power plant typically takes at least 4 years \cite{NuclearEnergyInstitute2015}.  Furthermore, nuclear reactors are incredibly difficult to hide, as seen by Israel's 2007 air raid against Syria \cite{WorldNuclearAssociation2015a}.  Thus, it seems unlikely that ISIL would be able to complete such a reactor. Therefore, there is a very low probability that ISIL could covertly produce it's own plutonium, and thus little need to monitor for reprocessing signals from ISIL territory.

@@

\subsection{???????? Scott's added brief RDD info, needs content proofing. Andrew tried to use it to link the above section to the case study below    ?????????}

@@

From the U.S. perspective, 9/11 serves as a case study to understand the shortcomings of IC, subsequent recommended changes, and possible shortcomings facing the IC with respect to ISIL. The 9/11 Report found that there are cultural and organizational factors that hinder strategic communication between U.S intelligence agencies (See \autoref{app:partners} for a disucussion on strategic communication on the international level, which also applies at the interagency level).  Prior to 9/11, U.S intelligence agencies operated in a culture dominated by \enquote{need-to-know}, which implied that organizations were given discretion in sharing data \cite{McConnell2008}. In the context of a dynamic ISIL threat, this means that agencies often have static analyses formed from an incomplete picture and lacking a competing analysis derived from the same data. The divisions between individual agencies render them incapable of accessing, prioritizing, translating, analyzing, and relaying the information for the policymakers to respond to in a timely manner. As a result, information flow is hindered by each agency's networks and databases lacking protocols to interact with other agencies' networks and databases.  Organizational inertia, standard operating procedures, and guidelines hinder the ability to change or implement new ideas. For example, the CIA's culture of risk aversion in the 1990s was detrimental to its intelligence capabilities and prevented them from properly assessing and taking action against the rising threat of terrorism \cite{Zegart2005}.

The situation is only complicated when considering international intelligence agencies and the sharing of data. To simplify the problem, the factors considered in determining the level of cooperation and the nature of an intelligence partnership between U.S. and other countries were distilled down to the historical relationships between both parties, the country's relationship to ISIL, its military and intelligence capability, and its vulnerability and geographical proximity to ISIL. The role of international entities is gauged by the organization's functions, credibility, and past success in relaying information and promptly responding to crises.  These factors are further described in \autoref{app:intel}.

@@

More than 60 countries have joined the \enquote{global coalition to degrade and defeat ISIL} \cite{Drennan2014}. The member countries can choose their method and level of participation. Saudi Arabia trains and equips Syrian opposition forces. Jordan's key role will be intelligence acquiring and sharing. Germany, France, and the United Kingdom are conducting airstrikes.

The number of countries in the coalition is growing with Obama's campaign, but the bar for inclusion is fairly low. While many countries have committed military capabilities to combatting ISIL, some countries join the coalition by merely declaring their intent to contribute. All these member countries don't share the same understanding of whom the real enemy is - ISIL or the Syrian regime led by President Bashar al-Assad. For example, Turkey has urged the coalition to topple the Assad regime as a condition to provide its air bases to carry out attacks against ISIL.

@@

UK and France have been U.S.' allies even since the country launched its global fight against terrorism after the 9/11 incident. Hence, the U.S. can expect UK specifically to be its closest ally in preventing ISIL's nuclear terrorism effort. The two countries \enquote{special relationship} dated back to Churchillian era means that the U.S. can usually count on the UK for mutual support, especially considering ISIL's threats to the country is very high. UK has been ISIL's main target for terrorist attacks, and given UK's nuclear capability, the country is extra vulnerable to terrorist attacks on any of its nuclear installations. So far, UK has joined U.S.-led coalition against ISIL, nevertheless, prospect for bilateral cooperation between the two countries cannot be disregarded.

Meanwhile, France is U.S.' oldest ally and relationship remains active and friendly. France supported U.S. in various issues, however, France was skeptical when U.S. invaded Iraq after the 9/11 incident. ISIL has also posed great threats to France (being a successful nuclear energy producer and all), France can be vulnerable to attacks on its nuclear energy facilities as well as its possession of nuclear weapons stocks. France is also among ISIL's main recruitments considering France high marginalized and discontented Muslim populations. France's high ISIL threats and vulnerabilities made the country a good collaborator in effectively fighting ISIL.

Nuclear capabilities in the hand of weak or unstable government give opportunity for ISIL to acquire nuclear capabilities. Nuclear capabilities of Pakistan and its weak governance as well as security vulnerabilities is of concern to the U.S.. Looking at U.S.-Pakistan relationship, it has its ups and down. Over the decades, Pakistan has been a great ally to the U.S., but there are a few instance where the two countries were in disagreement. U.S. viewed Pakistan as its ally and threat in the country's global war on terror, in which Pakistan is seen as the \enquote{weakest link} as it continue to fail in containing act of terrorism going on in its own country. U.S. need to prop Pakistan and reduce the country's vulnerabilities of ISIL penetrations and the possibility of ISIL acquiring Pakistan's nuclear weapons.

India is also another country with nuclear weapon capability, and is mostly on the \enquote{other side} of U.S.' security policies. India, together with Russia was among the pioneers of Non-Aligned Movement (NAM), in which views on security issues tend to be divergent with the U.S.. However, as of recently, the U.S. and India have set up \enquote{global strategic partnership}, based on shared democratic values and increasing convergence of interests on bilateral, regional and global issues; key recent developments in the relationship include the rapid growth of India's economy and bilateral trade, the close links between the Indian and American computer and Internet industries, a geopolitical coalition to balance the rise of China, the weakening of U.S.-Pakistan relations, among others. U.S. has cooperated with India in area of civil nuclear, and this can be expanded to development of nuclear security hopefully in the future. However, unlike Pakistan, India believed ISIL's threats as periphery issue to the country.

In analyzing Russia as a possible bilateral partner, the two country has had strained relationship, increasingly since the economic sanctions imposed by the U.S. to Russia. Dispute between the two countries have centered on power struggle for new world order, and recently exacerbated by Russia's annexation of Crimea. The two countries have been historical rival and the Cold War has proven to be one of the expansive and stretched out power struggle between the two. However, both countries realized and agreed to maintain dialogue in several areas including terrorism. Being misaligned with Russia on ISIL and nuclear terrorism issue could be serious mistakes as Russia's nuclear capability, if not protected and are allowed to be shared in the wrong hand, could have grave catastrophic impact to global security environment.

China, another U.S.' partner and rival at times, also have nuclear capabilities. China and the U.S. has close partnership particularly in economic sector, two-way trade between the two countries is over U.S.\$590 billion in 2014 alone. Chinese Muslim Uyghur militants have been associated with ISIL and efforts to undertake terrorist attacks in China. This strengthen the need for collaboration between China and U.S.; to ensure any ISIL activities can be contained in China and the country's nuclear weapons is protected from any probability of ISIL acquiring them.

@@

Iran and U.S. shared a common enemy factor in terms of ISIL. As the nuclear talks between the two countries are being hashed out, both countries have committed on fighting ISIL together. This can be a good opportunity for U.S. to combat ISIL more effectively as U.S. cannot disregard Iran's influence and role in Iraq's security landscape and their fight against ISIL.
Meanwhile, collaborating with Israel is considered a gray area and not clear-cut as it seems. Israel has been known to voice their support of ISIL, keeping in line with ISIL'ss common enemy with Israel i.e. Iran. As U.S. and Israel have been a strong ally in the Middle East region, this fact presented a conundrum to the U.S..

The U.S. may find other allies in the Middle East region, however, the region is deeply divided in their positions against the ISIL. Although most Middle Eastern countries sees ISIL as a serious threat to the region and its people, some still supported ISIL, if not openly but explicitly. U.S. has worked closely with Jordan, with the U.S. creating a U.S.\$5 billion counterterrorism partnership fund with Jordan and other countries (Libya, Lebanon, Syria) that are considered on the \enquote{frontlines} of U.S.' fights against ISIL.

The U.S. can find allies in European countries as well, as the region is facing \enquote{recruitment} threats from ISIL. Thousands of Europeans have been recruited by ISIL to fight jihad cause in ISIL's territories. ISIL is also setting up bases in a number of European countries, expanding its influence beyond the Middle East region. The U.S. can identify these countries and create cooperation efforts in order to effectively eliminate ISIL.

@@

Passive signals for nuclear materials include the gamma rays that are given off during natural radioactive decay of the material, and neutrons that result either from spontaneous fission or (\(\alpha\),n) nuclear reactions (the \(\alpha\) particle comes from natural \(\alpha\) decay). The two most likely and common types of material that could be used for a nuclear weapon are highly enrich \ce{^{235}U} and highly enriched \ce{^{239}Pu}, and so these materials will be considered for passive detection. Weapons grade Uranium has a very low rate of released neutrons, and so detection by gamma rays would be recommended. Weapons grade Plutonium releases both, a large amount of neutrons and gamma rays and so both methods of detection are possible.

@@

Of the radiation types mentioned: alpha, beta, and gamma, the gamma emitters are by far the farthest penetrating, most energetic, radioisotope emitters listed.  The shielding suggested will merely attenuate the energy levels of the gamma emissions, but in essence the emission passes through the shielding and continues on its travel through the medium. \autoref{fig:shielding_reqs} shows relative stopping distances in materials for given radiation types. Alphas, although they can be very energetic upon initial release from the source, are very large particles in comparison to the surroundings.  It is more likely that the 2+ charged alpha ion will interact with matter than the other radiation types, even as depicted a piece of paper is sufficient shielding material to attenuate alpha emissions.

@@

There are three approaches to disrupting ISIL'ss operations that will be particularly important in ensuring that ISIL does not acquire a nuclear device:

@@

ISIL's principal source of income is derived from control and sale of oil, which brings in about \$1 million a day [1]. Additional funds come from kidnap for ransom, extortion, criminal activities, and a small amount from external donations.

@@

A new report by Financial Action Task Force (an international body) finds that it is unclear whether ISIL's revenue collection will be sustainable over time, and that ISIL will have to gain more territory to remain financially viable \cite{TheEditorialBoard2015,Report2015}.

@@

In particular, sanctions will be levied against ISIL leadership, supporters, and financial facilitators. Effective use of sanctions against the CFO-like figures who manage ISIL funding networks will frustrate ISIL's ability to raise money and recruit fighters \cite{Cohen2014}.
